User-level Grid Monitoring with Inca 2

### 2007-02-22 pubs/papers/hpdc-gmw-2007/paper.tex
mods to arch intro and added reporter subsection
git-svn-id: file:///localdisk/subversion/inca/trunk/pubs@6538 7dba3f4a-8be6-0310-8b3b-b4fec25ea7f3

\documentclass[times, 10pt,twocolumn]{article} 
\usepackage{epsfig}
\usepackage{latex8}
\usepackage{times}
\usepackage{url}

@@

% What user-level monitoring provides and why we think it is good
User-level Grid monitoring approaches testing and performance measurement of
the Grid infrastructure from the user perspective.  The goal is to act as an
impartial test user of the Grid and detect problems with the infrastructure so
they can be fixed before users notice them -- user complaints should not be
the first indication of Grid failures.  We refer to an executable program that
tests or measures some aspect of a system or installed software from the
user-perspective as a reporter.  User-level Grid monitoring has the following
characteristics:

@@

\item Tests or performance measurements should be executed locally on all Grid
resources when appropriate so that all Grid access points available to users
are verified.  Similarly, it is important to executed some tests all-to-all
in order to detect site-to-site configuration errors such as authentication
problems.

@@

\noindent In our experience, writing and maintaining user-level tests and
performance measurements is an iterative refinement process since Grids are
dynamic and the software environment changes over time as packages are
upgraded.  It is also difficult to write and deploy a test perfectly the first
time because of portability issues and sometimes incomplete user
documentation.  For example, it may take a few iterations to determine whether
detected failures are the result of a faulty test, fault user documentation,
or a faulty system.  Realistically, this is a labor-intensive activity and
thus the amount of user-level Grid monitoring is often limited by the amount
of people resources that can be assiged to it.  However, A good goal is to
test all the commands that exist in the user documentation and/or have at
least one basic test for each software component of the Grid infrastructure.

% Inca 1~\cite{inca1} first addressed user-level Grid monitoring but had
% limitations.  We learned lessons from our TeraGrid deployment.  
In 2003, SDSC, in partnership with TeraGrid, began developing Inca
1~\cite{inca1} to implement a user-level monitoring system with the above
features and started using it to validate and
verify that CTSS was deployed consistently across all TeraGrid resources and
to monitor its status.  At that time, the only available user-level Grid
monitoring tools were the NCSA TestGrid script~\cite{ncsa-test} and
GITS~\cite{gits}, scripts that run a fixed number of Grid tests and output
results in HTML.  Although these tools were easy to install and produced
useful information, they showed only the view of the Grid from a single
resource, lacked automation, and were not easily extensible.  The initial
version of Inca was implemented as a client-server architecture.  It provided
data collection, data storage (with limited archiving capabilities) that was
accessible from a Web services interface, and data display though Web status
pages.  Inca was first deployed to TeraGrid in mid 2003 and released in late
2003. After running Inca 1 for a year and a half on TeraGrid, we learned some
valuable lessons and designed Inca 2, our current release.  Inca 2
contains a number of substantial improvements over Inca 1 with respect to
security, installation and maintenance, and storage capabilities.  A
production release of Inca 2 was provided in February 2007 and has been
running on TeraGrid since November 2006.  Currently, seven other Grids in the
U.S., Europe, and Australia use Inca.

@@

% Philosophy of user-level testing
% Practical byproducts

1.  Periodic tests executed by the monitoring system should be automated and
run from a standard user account.  A user GSI credential should be issued and
used by the monitoring system to test services that require authentication.  
%myproxy stuff
2.  Tests should be executed locally on all Grid resources when appropriate so
that all Grid access points available to users are tested.  
3.  The configuration of tests should be managed centrally in order to ensure
consistent testing of the resources, rather than by system administrators who
may not use the same information that is available to users.
4.  The testing system should easily adapt to new resources and testing
requirements.  In our experience, testing has an iterative refinement process
because Grids are dynamic and the software environment changes over time as
packages are upgraded.  Furthermore, it is often difficult to write and deploy
a test perfectly the first time because of portability issues and incomplete
user documentation.   
5.  System administrators should not need to know the details of the
monitoring system in order to troubleshoot problems.  When a problem is
detected the monitoring system should provide system administrators with
sufficient information to locate the problem easily and test repairs.
6.  A history of test results (especially any error messages) should be
retained in order to understand the behavior of their Grid over time. This
history should be available through a flexible querying interface.
7.  A running monitoring system should have a limited impact on the resources
it measures and should be secure.

%------------------------------------------------------------------------- 
\Section{Architecture}

\SubSection{Overview}
% may not need to leave in here -- intro is assumed to be overview
% Overview of all architecture components -- arch picture
% move text to here I think
1) The user creates a suite using incat and submits it to the agent

2) The agent 
    a) fetches reporters from the reporter repository
    b) creates a reporter manager on each resource
    c) sends the suite and reporters to each reporter manager.

3) Each reporter manager executes reporters according to its schedule and
sends results (reports) to the depot

4) Data consumers display collected data (reports) by querying the depot

\begin{figure*}[Htb]
  \centering
  \mbox{\epsfig{file=arch.eps, width=.8\textwidth}}
  \caption{\label{arch_fig} Inca architecture.}
\end{figure*}

@@

To configure reporter execution, an Inca administrator launches a GUI tool,
called incat.  Incat allows the administrator to choose which resources to
monitor and which reporters to deploy to those resources.  For each
reporter, the administrator can specify the resources to run on,
command-line arguments, the runtime environment, the frequency of execution,
and resource limits. The Inca administrator may also group related reporters
into suites that can be shared across different Inca deployments.  This can
be useful to determine interoperability among Grids or to determine whether
an application's requirements are being fulfilled on a Grid.

\subsubsection{Reporters}

Each Inca reporter is an executable program that tests or measures some
aspect of a system or installed software.   Reporter executables are
designed to be easy to produce and can be run outside of the Inca system
(e.g., by a system administrator).  A reporter could be a simple Globus
gatekeeper ping test (see Figure 2) or a more complex Grid application
benchmark.  Reporters must support certain command line options and produce
XML according to the Inca reporter schema.  The reporter schema requires a
header describing the context of the reporter execution (e.g., reporter
name, arguments, working directory, timestamp) and a body containing the
results expressed as any XML sequence. This flexible body schema enables
reporters to express a wide variety of information.  Today, we have three
standard body schemas to express software version information, software
functionality or service tests results, and usage information.  We provide
Perl APIs to handle much of the effort of writing reporters.  Most current
reporters use these APIs and consist of fewer than 30 lines of code.  
% capture context of execution

\subsubsection{Reporter Repositories}

>>>>

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{epsfig}
\usepackage{latex8}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{times}
\usepackage{url}

@@

% What user-level monitoring provides and why we think it is good
User-level Grid monitoring approaches testing and performance measurement of
the Grid infrastructure from the user perspective.  The goal is to act as an
impartial test user of the Grid and detect problems with the infrastructure so
they can be fixed before users notice them -- user complaints should not be
the first indication of Grid failures.  We refer to an executable program that
tests or measures some aspect of a system or installed software from the
user-perspective as a reporter.  In our view, the following guidelines 
define user-level Grid monitoring:

@@

\item Tests or performance measurements should be executed locally on all Grid
resources when appropriate so that all Grid access points available to users
are verified.  Similarly, it is important to execute some tests all-to-all
in order to detect site-to-site configuration errors such as authentication
problems.

@@

\noindent In our experience, writing and maintaining user-level tests and
performance measurements is an iterative refinement process since Grids are
dynamic and the software environment changes over time as packages are
upgraded.  It is also difficult to write and deploy a test perfectly the first
time because of portability issues and sometimes incomplete user
documentation.  For example, it may take a few iterations to determine whether
detected failures are the result of a faulty test, fault user documentation,
or a faulty system.  Realistically, this is a labor-intensive activity and
thus the amount of user-level Grid monitoring is often limited by the amount
of people resources that can be assigned to it.  However, A good goal is to
test all the commands that exist in the user documentation and/or have at
least one basic test for each software component of the Grid infrastructure.

% Inca 1~\cite{inca1} first addressed user-level Grid monitoring but had
% limitations.  We learned lessons from our TeraGrid deployment and
% developed Inca 2.
In 2003, SDSC, in partnership with TeraGrid, began developing Inca
1~\cite{inca1} to implement a user-level monitoring system with the above
features and started using it to validate and
verify that CTSS was deployed consistently across all TeraGrid resources and
to monitor its status.  At that time, the only available user-level Grid
monitoring tools were the NCSA TestGrid script~\cite{ncsa-test} and
GITS~\cite{gits}, scripts that run a fixed number of Grid tests and output
results in HTML.  Although these tools were easy to install and produced
useful information, they showed only the view of the Grid from a single
resource, lacked automation, and were not easily extensible.  The initial
version of Inca was implemented as a client-server architecture.  It provided
data collection, data storage (with limited archiving capabilities) that was
accessible from a Web services interface, and data display though Web status
pages.  Inca was first deployed to TeraGrid in mid 2003 and released in late
2003. After running Inca 1 for a year and a half on TeraGrid, we learned some
valuable lessons and designed Inca 2, our current release.  Inca 2
contains a number of substantial improvements over Inca 1 with respect to
security, installation and maintenance, and storage capabilities.  A
production release of Inca 2 was provided in February 2007 and has been
running on TeraGrid since November 2006.  Currently, seven other Grids in the
U.S., Europe, and Australia use Inca.

@@

% Philosophy of user-level testing
% Practical byproducts
The goal of the Inca 2 design was to

\begin{enumerate}
\item A reporter 
\item logging in reporters
\item Given the challenges of writing portable tests and benchmarks, Inca 2 
includes a set of Perl APIs that provide convenience functions to simplify
this process.

\item Periodic tests executed by the monitoring system should be automated and
run from a standard user account.  A user GSI credential should be issued and
used by the monitoring system to test services that require authentication.  
%myproxy stuff
\item Tests should be executed locally on all Grid resources when appropriate so
that all Grid access points available to users are tested.  
\item The configuration of tests should be managed centrally in order to ensure
consistent testing of the resources, rather than by system administrators who
may not use the same information that is available to users.
\item The testing system should easily adapt to new resources and testing
requirements.  In our experience, testing has an iterative refinement process
because Grids are dynamic and the software environment changes over time as
packages are upgraded.  Furthermore, it is often difficult to write and deploy
a test perfectly the first time because of portability issues and incomplete
user documentation.   
\item System administrators should not need to know the details of the
monitoring system in order to troubleshoot problems.  When a problem is
detected the monitoring system should provide system administrators with
sufficient information to locate the problem easily and test repairs.
\item A history of test results (especially any error messages) should be
retained in order to understand the behavior of their Grid over time. This
history should be available through a flexible querying interface.
\item A running monitoring system should have a limited impact on the resources
it measures and should be secure.
\end{enumerate}

%------------------------------------------------------------------------- 
\Section{Inca 2 Architecture}

\begin{figure*}[htb]
  \centering
  \mbox{\epsfig{file=arch.eps, width=.8\textwidth}}
  \caption{\label{arch_fig} Inca architecture.}
\end{figure*}

% overview description of components
Figure~\ref{arch_fig} shows the architecture of Inca 2 designed to implement
the features described in the previous section.  Inca 2 is composed of 3 core
components (in highlighted box):  the agent, depot, and reporter manager.
The \textit{agent} and \textit{reporter managers} coordinate the execution of
tests and performance measurements on the Grid resources and the
\textit{depot} stores and archives the results.  The inputs to Inca 2 are one
or more \textit{reporter repositories} that contain user-level tests and
benchmarks called \textit{reporters} and a configuration file describing how
to execute them on the Grid resources created using an administration GUI tool
called \textit{incat} (Inca Administration Tool).  The output or results
collected from the resources are queried by the \textit{data consumer} and
displayed to users.  The following steps describe how an Inca administrator
would deploy user-level test and/or performance measurements to their
resources.

\begin{enumerate}

\item Using the guidelines described in the introduction, the Inca
administrator either writes reporters to monitor their Grid or uses existing
reporters in a published repository.

\item The Inca administrator creates a deployment configuration file that
describes the user-level monitoring for their Grid using incat and submits it
to the agent.

\item The agent
  \begin{enumerate}
    \item fetches reporters from the reporter repository
    \item creates a reporter manager on each resource
    \item sends the reporters and instruction for executed them to each reporter manager.
  \end{enumerate}

\item Each reporter manager executes reporters according to its schedule and
sends results (reports) to the depot

\item Data consumers display collected data (reports) by querying the depot
\end{enumerate}

\noindent The following subsections describe the Inca components in more
detail using the order of the steps above.

\SubSection{Reporters}

\lstset{
  basicstyle=\scriptsize\ttfamily, 
  frame=single,
  keywordstyle=\textbf, 
  identifierstyle=, 
  commentstyle=\scriptsize\ttfamily, 
  stringstyle=\ttfamily, 
  numbers=left, 
  numberstyle=\scriptsize,
  stepnumber=2,
  firstnumber=1,
  showstringspaces=false} 
\begin{figure}
\lstset{language=Perl} 
\subfigure[]{\lstinputlisting{grid.globus.gramPing}}
\lstset{language=XML,keywordstyle={}} 
\subfigure[]{\lstinputlisting{grid.globus.gramPing.out}}
\caption{\label{pingReporter}(a) An Inca reporter that tests the user-level
availability of a Globus GRAM server and (b) its sample output.}
\end{figure}

An Inca reporter is an executable program that tests or measures some aspect
of a system or installed software.   Reporter executables are designed to be
easy to produce and can be run outside of the Inca system (e.g., by a system
administrator).  A reporter should be written following the user-level
monitoring guidelines described in the introduction and could be a simple
Globus~\cite{globus} GRAM gatekeeper ping test (see Figure~\ref{pingReporter})
or a more complex Grid application benchmark~\cite{grasp}.  Reporters must
support certain command line options and produce an XML report according to
the Inca reporter schema.  The goal of the reporter schema is to accomodate
multiple types of data and to capture information about the the reporter
execution to diagnose any detected failures.  The reporter schema consists of
the following elements:

\begin{itemize}
\item a set of header tags describing the context of the reporter
execution:  GMT timestamp, hostname, reporter name, reporter version, working
directory, reporter path, and arguments;
\item an optional set of debug or information log tags similar to log4j~\cite{log4j} output;
\item a body containing the results expressed as any XML sequence; and
\item an exit status tag
indicating whether the reporter was able to complete its test or measurement
and an optional error message.  
\end{itemize} 
\noindent Because the body of the report can be any XML sequence, it enables
reporters to express a wide variety of information.  Today, we have four
standard body schemas to express software version information, software
functionality or service tests results, usage information, and performance
results.

We provide an extensible set of Perl APIs to handle much of the effort of
writing reporters.  Figure~\ref{pingReporter} shows (a) an Inca reporter that
pings a pre-WS Globus GRAM server and (b) its corresponding output when
executed on a local machine.  The reporter uses two of the Perl APIs:
Inca::Reporter::SimpleUnit (which is a subclass of Inca::Reporter) and
Inca::Reporter::GridProxy.  The Inca::Reporter API provides convenience
functions for handling the required input arguments
(lines~\ref{pingReporter}(a) 12-14 $\rightarrow$ lines \ref{pingReporter}(b)
10-31) and printing the Inca-compliant header and exit status XML tags
(lines\ref{pingReporter}(a) 4-11, 24 $\rightarrow$ \ref{pingReporter}(b)
1-9,43-46).  It also includes some informational and debug logging functions
again similar to log4j; the log output has been very useful to system
administrators who can view a summary of the test or benchmark that the
reporter performed without reading the reporter source code.  Line
\ref{pingReporter}(a) 15 shows a special log function, loggedCommand, that
executes a system command (using a 30 second timeout) and logs to the 'system'
level of the log XML output (lines \ref{pingReporter}(b) 32-36).  The subclass
Inca::Reporter::SimpleUnit is the API responsible for printing the simplist
standard reporter body schema -- reporting software functionality or service
tests results; it handles the printing of the small body XML
(lines~\ref{pingReporter}(b) 39-41) using the functions unitSuccesss
(line~\ref{pingReporter}(a) 22) or unitFailure (lines~\ref{pingReporter}(a)
18, 20) depending on the output of system command.  And finally, the
Inca::Reporter::GridProxy API (line~\ref{pingReporter}(a) 3) adds a proxy
dependency to the reporter which tells the Inca system to download a
short-term proxy before running this reporter (this is discussed further in
Section~\ref{proxy}).  Most current reporters use the Perl APIs and consist of
fewer than 30 lines of code.

\SubSection{Reporter Repositories}

@@

The Inca administrative GUI tool (incat) is the main interface by which an
Inca administrator interacts with the Inca system.  Incat allows the Inca
administrator to view the choose which resources to monitor and which tests or 
benchmarks to deploy to those resources.  The following subsections
describe the Inca components that are configurable via incat.  
To configure reporter execution, an Inca administrator launches a GUI tool,
called incat.  Incat allows the administrator to choose which resources to
monitor and which reporters to deploy to those resources.  For each
reporter, the administrator can specify the resources to run on,
command-line arguments, the runtime environment, the frequency of execution,
and resource limits. The Inca administrator may also group related reporters
into suites that can be shared across different Inca deployments.  This can
be useful to determine interoperability among Grids or to determine whether
an application's requirements are being fulfilled on a Grid.
